### 2023-05-11 response-to-audit-report.md

# Response to Audit Report for the Marlowe Specification and for Marlowe on Cardano

## Executive Summary

Marlowe is a domain-specific language (DSL) for multi-party financial contracts. Operational semantics are formalized in the Isabelle proof-assistant and an implementation of Marlowe for the Cardano blockchain is written in the Haskell/Plutus programming language. Early in 2023 the Marlowe specification, the on-chain implementation and the validator test-suite were audited by Tweag's High Assurance Software Group. The report is available online. It categorizes findings of either high, medium or low severity. The present document summarizes the subsequent changes made to the Marlowe specification, the implementation of the validator and the property-based tests for the validator. All of the high-severity findings have been remedied or mitigated through changes to the specification, the proofs in Isabelle or the implementation. Nearly all of the medium-severity and low-severity findings have also been addressed; justification for not addressing the remainder has been provided.

The high-severity findings comprise handling of negative deposits, prevention of "double satisfaction", enforcement of state invariants, an implementation difference between the formal specification versus the Plutus implementation and the proof of money preservation in Isabelle. The Marlowe validator has been altered to resolve these first three findings so that it correctly prevents negative deposits from altering the balance of internal accounts, it only allows other Plutus scripts to run during transactions that do not make Marlowe payments, and it rigorously enforces invariants for initial and final states. The fourth finding regarding the Plutus implementation was mitigated via code analysis and property-based testing. The fifth finding about the money-preservation proof was remedied by revising the Isabelle proof.

The medium-severity findings typically suggested improvements to the Isabelle proofs, addition of more extensive property-based testing, correction of infelicities in the implementation, or safe-coding improvements. The low-severity findings generally centered on the usefulness of code comments, naming of variable bindings, omissions in specifications and correction of typographical errors. All of the findings that affect the clarity of the specifications, the coverage of testing, or the robustness of implementation were made. Numerous additional property-based tests were added to the test suite as a result of audit findings.

The changes outlined here have not been re-audited. The alterations to the Marlowe validator (detailed in the Appendix) have been kept to the minimum needed to address exclusively the audit findings.

Finally, please note that the scope of the audit included the Marlowe language and its Plutus interpreter, but not individual Marlowe contracts or off-chain code. Although the Marlowe language and its implementation on Cardano aim to make smart contracts safer, it is nevertheless still possible to design individual instances of Marlowe contracts that exhibit unwanted behavior. Such undesirable behavior may result either from poor contract design or lack of testing and analysis of the contract. [A best-practices guide](https://github.com/input-output-hk/marlowe-cardano/blob/main/marlowe/best-practices.md) and [a security guide](https://github.com/input-output-hk/marlowe-cardano/blob/main/marlowe/security.md) provide advice for the safe and secure design and testing of individual Marlowe contracts, which can themselves be subject to audit or certification.

@@

## Overview of Isabelle Changes

The audit report's main and high-severity concerns for the Isabelle Proofs were:

- 2.1.6 Missing description of Merkleization
- 2.1.4 Inaccurate formulation of Money preservation

We did not mention merkleization in the specification document as it was considered an implementation detail. This does lessen the applicability of the Isabelle proofs to the different implementations (Marlowe on Cardano, Marlowe in PureScript, etc). To check that an implementation is compliant with the specification we use the `spec-test` tool, which uses property based tests to check that a custom implementation behaves equally to the Isabelle implementation. Without the proper formalization, the tool can only test for the non-merkleized cases. In practice this gives us confidence that the merkleized interpreter works the same for non-merkleized cases. To test the correctness of merkleized cases, the marlowe-cardano repository has specific tests. The formalization of merkleization will be introduced in a future version of the specification (initial work started in [PR-171](https://github.com/input-output-hk/marlowe/pull/171).

@@

The Money preservation theory had only taken into account the amount of assets and not the type. This meant that the proof was not considering a case where a contract could receive `20 ada`,  `15 djed`, and pay back `20 djed` and `15 ada`. We added a new `MultiAssets` type and refactored the Isabelle code (without modifying the interpreter) to prove that the Assets are preserved.

Other medium and low priorities findings regarding documentation and legibility of the specification were also addressed.

## Overview of Changes to Marlowe-Cardano Specification and Haskell (Plutus) Implementation

The audit report's main and high-severity concerns for the Plutus validators for Marlowe on Cardano fell into four clusters:

@@

We renamed as suggested in [PR-182](https://github.com/input-output-hk/marlowe/pull/182).

### 2.2.11 Inaccurate specification of `giveMoney` *(Severity: Low)*

> **File `specification-v3-rc1.pdf`, Section `2.2.9 Utilities`, page *22***
>
> It says
>
> > The *giveMoney* function transfers funds internally between accounts.
>
> which is not accurate. It should say instead
>
> > The *giveMoney* function deposits funds to an internal account.
>
> This function is confusing in that it takes the account identifier of the paying account which is not used for anything other than filling a field in the returned value.

This function is confusing, we will likely remove it in a future release.

@@

> **File `specification-v3-rc1.pdf`, Section `3.1 Money preservation`, page *29***
>
> Money preservation is expressed with an equality. This equality, however, only ensures money preservation for those lists of transactions that produce no error. In other words, there is no guarantee that money will be preserved for those lists of transactions that fail.
>
> This is not a concern in practice because the lists of transactions that fail to evaluate are not accepted in the blockchain. However, this should be made explicit in the explanation of the property.

This was taken into account in the new Asset preservation theory and a [paragraph](https://github.com/input-output-hk/marlowe/pull/161/files#diff-060b0c9f7bde4e9ebc4e7b9d6a2202af75213be8d44d7f4745e02054be5186deR52-R53) was added to explain the case.

### 2.2.23 Complicated definition of `allAccountsPositive` *(Severity: Low)*

> **File `specification-v3-rc1.pdf`, Section `3.3 Positive Accounts`, page *30***
>
> The definition of\ `allAccountsPositive` is complicated and can be refactored as `all ((_, money) -> money > 0)`.

### 2023-05-11 response-to-audit-report.md

# Response to Audit Report for the Marlowe Specification and for Marlowe on Cardano

## Executive Summary

Marlowe is a domain-specific language (DSL) for multi-party financial contracts. Operational semantics are formalized in the Isabelle proof-assistant and an interpreter for Marlowe for the Cardano blockchain is provided as Plutus validator script. Early in 2023 the Marlowe specification, the on-chain implementation and the validator test-suite were audited by Tweag's High Assurance Software Group. The report is available online. It categorizes findings of either high, medium or low severity. The present document summarizes the subsequent changes made to the Marlowe specification, the implementation of the validator and the property-based tests for the validator. All of the high-severity findings have been remedied or mitigated through changes to the specification, the proofs in Isabelle or the implementation. Nearly all of the medium-severity and low-severity findings have also been addressed; justification for not addressing the remainder has been provided.

The high-severity findings comprise handling of negative deposits, prevention of "double satisfaction", enforcement of state invariants, an implementation difference between the formal specification versus the Plutus implementation and the proof of money preservation in Isabelle. The Marlowe validator has been altered to resolve these first three findings so that it correctly prevents negative deposits from altering the balance of internal accounts, it only allows other Plutus scripts to run during transactions that do not make Marlowe payments, and it rigorously enforces invariants for initial and final states. The fourth finding regarding the Plutus implementation was mitigated via code analysis and property-based testing. The fifth finding about the proof of the money preservation theorem was remedied by revising the Isabelle code.

### 2023-05-11 response-to-audit-report.md

# Response to Audit Report for the Marlowe Specification and for Marlowe on Cardano

## Executive Summary

Marlowe is a domain-specific language (DSL) for multi-party financial contracts. Operational semantics are formalized in the Isabelle proof-assistant and an interpreter for Marlowe for the Cardano blockchain is provided as Plutus validator script. Early in 2023 the Marlowe specification, the on-chain implementation and the validator test-suite were audited by Tweag's High Assurance Software Group. The report is available online. It categorizes findings of either high, medium or low severity. The present document summarizes the subsequent changes made to the Marlowe specification, the implementation of the validator and the property-based tests for the validator. All of the high-severity findings have been remedied or mitigated through changes to the specification, the proofs in Isabelle or the implementation. Nearly all of the medium-severity and low-severity findings have also been addressed; justification for not addressing the remainder has been provided.

The high-severity findings comprise handling of negative deposits, prevention of "double satisfaction", enforcement of state invariants, an implementation difference between the formal specification versus the Plutus implementation and the proof of money preservation in Isabelle. The Marlowe validator has been altered to resolve these first three findings so that it correctly prevents negative deposits from altering the balance of internal accounts, it only allows other Plutus scripts to run during transactions that do not make Marlowe payments and it rigorously enforces invariants for initial and final states. The fourth finding regarding the Plutus implementation was mitigated via code analysis and property-based testing. The fifth finding about the proof of the money preservation theorem was remedied by revising the Isabelle code.

The medium-severity findings typically suggested improvements to the Isabelle proofs, addition of more extensive property-based testing, correction of infelicities in the implementation, or safe-coding improvements. The low-severity findings generally centered on the usefulness of code comments, naming of variable bindings, omissions in specifications and correction of typographical errors. All of the findings that affect the clarity of the specifications, the coverage of testing, or the robustness of implementation were made. Numerous additional property-based tests were added to the test suite as a result of the audit findings.

@@

The changes outlined here have not been re-audited. The alterations to the Marlowe validator (detailed in the Appendix) have been kept to the minimum needed to address exclusively the audit findings.

Finally, please note that the scope of the audit included the Marlowe language and interpreter, but not individual Marlowe contracts or off-chain code. Although the Marlowe language and its implementation on Cardano aim to make smart contracts safer, it is nevertheless still possible to design individual instances of Marlowe contracts that exhibit unwanted behavior. Such undesirable behavior may result either from poor contract design or lack of testing and analysis of the contract. [A best-practices guide](https://github.com/input-output-hk/marlowe-cardano/blob/main/marlowe/best-practices.md) and [a security guide](https://github.com/input-output-hk/marlowe-cardano/blob/main/marlowe/security.md) provide advice for the safe and secure design and testing of individual Marlowe contracts, which can themselves be subject to audit or certification.

@@

## Acknowledgement

We thank the High Assurance Software Group at Tweag for their diligent and expert analyses of the Marlowe specifications, proofs, implementations and tests and their clear and thorough communication of their findings.

@@

The audit report's main and high-severity concerns for the Isabelle Proofs were:

- 2.1.6 Missing description of Merkleization
- 2.1.4 Inaccurate formulation of Money preservation

We did not mention merkleization in the specification document as it was considered an implementation detail. This does lessen the applicability of the Isabelle proofs to the different implementations (Marlowe on Cardano, Marlowe in PureScript, etc). To check that an implementation is compliant with the specification we use the `spec-test` tool, which uses property based tests to check that a custom implementation behaves equally to the Isabelle implementation. Without the proper formalization, the tool can only test for the non-merkleized cases. In practice this gives us confidence that the merkleized interpreter works the same for non-merkleized cases. To test the correctness of merkleized cases, the marlowe-cardano repository has specific tests. The formalization of merkleization will be introduced in a future version of the specification (initial work started in [PR-171](https://github.com/input-output-hk/marlowe/pull/171).

The Money preservation theory had only taken into account the amount of assets and not the type. This meant that the proof was not considering a case where a contract could receive `20 ada`, `15 djed` and pay back `20 djed` and `15 ada`. We added a new `MultiAssets` type and refactored the Isabelle code (without modifying the interpreter) to prove that the Assets are preserved.

### 2023-05-11 response-to-audit-report.md

# Response to Audit Report for the Marlowe Specification and for Marlowe on Cardano

## Executive Summary

Marlowe is a domain-specific language (DSL) for multi-party financial contracts. Operational semantics are formalized in the Isabelle proof-assistant and an interpreter for Marlowe for the Cardano blockchain is provided as Plutus validator script. Early in 2023 the Marlowe specification, the on-chain implementation and the validator test-suite were audited by Tweag's High Assurance Software Group. The report is available online. It categorizes findings of either high, medium or low severity. The present document summarizes the subsequent changes made to the Marlowe specification, the implementation of the validator and the property-based tests for the validator. All of the high-severity findings have been remedied or mitigated through changes to the specification, the proofs in Isabelle or the implementation. Nearly all of the medium-severity and low-severity findings have also been addressed, otherwise justification for not addressing a finding has been provided.

@@

## Overview of Isabelle Changes

The audit report's main and high-severity concerns for the Isabelle Proofs were:

- 2.1.6 Missing description of Merkleization
- 2.1.4 Inaccurate formulation of Money preservation

We did not mention merkleization in the specification document as it was considered an implementation detail. This does lessen the applicability of the Isabelle proofs to the different implementations (Marlowe on Cardano, Marlowe in PureScript, etc). To check that an implementation is compliant with the specification we use the `spec-test` tool, which uses property based tests to check that a custom implementation behaves equally to the Isabelle implementation. Without the proper formalization, the tool can only test for the non-merkleized cases. In practice this gives us confidence that the merkleized interpreter works the same for non-merkleized cases. To test the correctness of merkleized cases, the marlowe-cardano repository has specific tests. The formalization of merkleization will be introduced in a future version of the specification (initial work started in [PR-171](https://github.com/input-output-hk/marlowe/pull/171)).

### 2023-05-11 response-to-audit-report.md

## Executive Summary

Marlowe is a domain-specific language (DSL) for multi-party financial contracts. Operational semantics are formalized in the Isabelle proof-assistant and an interpreter for Marlowe for the Cardano blockchain is provided as Plutus validator script. Early in 2023 the Marlowe specification, the on-chain implementation and the validator test-suite were audited by Tweag's High Assurance Software Group. The report is available online. It categorizes findings of either high, medium or low severity. The present document summarizes the subsequent changes made to the Marlowe specification, the implementation of the validator and the property-based tests for the validator. All of the high-severity findings have been remedied or mitigated through changes to the specification, the proofs in Isabelle or the implementation. Nearly all of the medium-severity and low-severity findings have also been addressed, otherwise justification for not addressing a finding has been provided.

The high-severity findings comprise handling of negative deposits, prevention of "double satisfaction", enforcement of state invariants, an implementation difference between the formal specification versus the Plutus implementation and the proof of money preservation in Isabelle. The Marlowe validator has been altered to resolve these first three findings so that it correctly prevents negative deposits from altering the balance of internal accounts, it only allows other Plutus scripts to run during transactions that do not make Marlowe payments and it rigorously enforces invariants for initial and final states. The fourth finding regarding the Plutus implementation was mitigated via code analysis and property-based testing. The fifth finding about the proof of the money preservation theorem was remedied by revising the Isabelle code.

The medium-severity findings typically suggested improvements to the Isabelle proofs, addition of more extensive property-based testing, correction of infelicities in the implementation or safe-coding improvements. The low-severity findings generally centered on the usefulness of code comments, naming of variable bindings, omissions in specifications and correction of typographical errors. All of the findings that affect the clarity of the specifications, the coverage of testing or the robustness of implementation were made. Numerous additional property-based tests were added to the test suite as a result of the audit findings.

@@

The changes outlined here have not been re-audited. The alterations to the Marlowe validator (detailed in the Appendix) have been kept to the minimum needed to address exclusively the audit findings.

Finally, please note that the scope of the audit included the Marlowe language and interpreter, but not individual Marlowe contracts or off-chain code. Although the Marlowe language and its implementation on Cardano aim to make smart contracts safer, it is nevertheless still possible to design individual instances of Marlowe contracts that exhibit unwanted behavior. Such undesirable behavior may result either from poor contract design or lack of testing and analysis of the contract. [A best-practices guide](https://github.com/input-output-hk/marlowe-cardano/blob/main/marlowe/best-practices.md) and [a security guide](https://github.com/input-output-hk/marlowe-cardano/blob/main/marlowe/security.md) provide advice for the safe and secure design and testing of individual Marlowe contracts which can themselves be subject to audit or certification.
